---
title: "Class 11: Population Analysis HW"
author: "Kaitlyn Madriaga, A17217752"
format: pdf
editor: visual
---

## Population Scale Analysis 


```{r}
file <- read.table(url("https://bioboot.github.io/bggn213_W19/class-material/rs8067378_ENSG00000172057.6.txt"))

head(file)
```
> Q13. Read this file into R and determine the sample size for each genotype and their
corresponding median expression levels for each of these genotypes

```{r}
summary(file)
table(file$geno)

```
There are 462 samples-- 108 A/A, 233 A/G, 121 G/G. 

```{r}
box <- boxplot(file$exp ~ file$geno, xlab = "Genotype", ylab = "Expression")
box$stats
```
The median for A/A is 31.2487, the median for A/G is 25.06486, and the median for G/G is 20.07363.  

> Q14.  Generate a boxplot with a box per genotype, what could you infer from the relative
expression value between A/A and G/G displayed in this plot? Does the SNP effect the
expression of ORMDL3?

```{r}
library(ggplot2)
```

```{r}
ggplot(file) + aes(geno, exp, fill = geno) +
  geom_boxplot(notch = TRUE)
```
The expression values of A|A and G|G are very different, as their medians/notches do not overlap each other.  We can infer that with G|G genotype have a lower expression of the ORMDL3.   
